---
title: "ESM 206 Lab 6"
author: "Julia Wilson"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Example of a rank-based test 

We'll make our own samples, using a pseudorandom generator. 

Notes: 
telling it what you want to maximum value to be 
random sample of values between 0 - 20 
set.seed allows us to all have the same random sample, good for reproducibility
set.seed(the number here is where you are starting from); 
- tell r to create a random sample of integer values, there's a huge sequence of values where it could start from, where the starting point is. Settting the seed allows us to say "if you are oulling form this giant sequence, I want you to start at 1414 or 2." Starting point of values from where vector is pulled. 
replace = once a value of observatio is removed, the observation value is put back into the pool. This is why there are multiple 7s, 17s etc. 
- pseudo random bc once you set seed it's not truly random \
- try not to use a single random sample, but a bunch of random samples together 
- the sample. int(#) is the max number your sequence can go to 

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Create quick histrograms. These shows not a normal distribution in gp_1. gp_2 is uniform, but not necessarily normally distributed. 

```{r}
hist(gp_1)

hist(gp_2)
```

Try a t-test: 
```{r}
t.test(gp_1, gp_2)
```
Meaning of p value: The p value indicates that if these samples were draw from population with the same mean, then there is a 19.8% chance of taking two random that the samples that are at least as different as the sample means we found by random chance. 

If these samples were drawn from populations with the same mean, the prob of taking two random samples with means *at least as different* as the sample means we found by random chance (taking into account spread, n) is 19.8%

Super likely null hypothesis is true and that we will get these 

Retain or fail to reject the null hypothesis >> There is no significant difference in means between group 1 and group 2. 

Warning: people get weirdly upset if you say "accept" the null 

Now let's compare this outcome to a rank-based test. 

## Mann Whitney U unparied rank-based test 

often recorded as a comparison of medians 
if we have matching values, then you get tie in the ranks and then there is just an approximation made when you do the calculation 

```{r}
mwu <- wilcox.test(gp_1, gp_2)

mwu 
```

If these samples were drawn from populations with the same ranks (medians), the
probability of taking two random samples with ranks at least as different as the
sample ranks we found by random chance is 0.28. 

Is that so unlikely to have happened that my difference in rank of my samples, that i should reject my hypothesis? 

28% of the time is very likely that these samples are drawn from populations with the same ranks or medians. 

There is no significant difference in ranks between group 1 and group 2 (statistical summary).

Median scores for group 1 (M = 14) and group 2 (M = 12) did not differ significantly (Mann Whitney U test: U(df) = 86, p = 0.28).

If you chose what the appropriate type of stats hypothesis, then every other metric should align. Don't report median as central tendancy and then use t test using a mean. Make sure what appropriate test to use that it aligns with everything else you are presenting. Also true for data visualization. 













